---
title: "AK_Ecopath"
author: "Bia Dias"
date: "3/30/2022"
output: 
  powerpoint_presentation:
  widescreen: true
---

```{r setup, include=FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = FALSE)

#url <- "https://download.r-forge.r-project.org/bin/windows/contrib/4.4/rgdal_1.6-7.zip"
#install.packages(url, type="source", repos=NULL)

library(rgdal)
library(rbgm)
library(sf)
library(ggplot2)
library(here)
library(tidyverse)
library(ggrepel)
library(ggpattern)


```

## AK ecosystem modeling working group

Modeling group for Ecopath and Atlantis 

## Model and data domains

- Alaska Marine Areas from the ecosystem status report - Bridget
- Atlantis model - Alberto
- Prince William Sound - Bia
- Central Gulf of Alaska - Bia
- Chukchi Sea - Andy
- Eastern Bering Sea - Andy
- Western GOA - Bridget
- Eastern GOA - Szymon



```{r AK_ESR, echo = FALSE, warning = FALSE, include=FALSE}

#Alaska polygon
AKpoly <- read_sf("shapefiles/AKpoly.shp")
AKpoly <- st_transform(x = AKpoly, "+proj=aea +lat_0=50 +lon_0=-154 +lat_1=55 +lat_2=65 +x_0=0 +y_0=0 +ellps=GRS80 +units=m +no_defs")

fgdb <- "AK_Crab_Management_Areas.gdb"

# List all feature classes in a file geodatabase
subset(ogrDrivers(), grepl("GDB", name))
fc_list <- ogrListLayers(fgdb)

fc <- readOGR(dsn=fgdb,layer="Alaska_Marine_Areas_AK_prj")

fc2 <- spTransform(x = fc, CRS("+proj=aea +lat_0=50 +lon_0=-154 +lat_1=55 +lat_2=65 +x_0=0 +y_0=0 +ellps=GRS80 +units=m +no_defs"))

AKmngtArea <- st_as_sf(fc2)
AKmngtA <- st_transform(AKmngtArea, crs = "+proj=aea +lat_0=50 +lon_0=-154 +lat_1=55 +lat_2=65 +x_0=0 +y_0=0 +ellps=GRS80 +units=m +no_defs")

#All the information from ESR


```

## Data domains

```{r AK_ESR2, warning = FALSE, out.width="90%"}
plot(AKmngtArea)
```


```{r Atlantis, echo=FALSE}

## Atlantis model ####
atlantis_bgm <- read_bgm('GOA_WGS84_V4_final.bgm') #Albi's extent 
atlantis_box <- box_sf(atlantis_bgm) # turn bgm to sf
atlantis_crs <- atlantis_bgm$extra$projection # custom CRS

atlantis_box2 <- st_transform(atlantis_box, crs = "+proj=aea +lat_0=50 +lon_0=-154 +lat_1=55 +lat_2=65 +x_0=0 +y_0=0 +ellps=GRS80 +units=m +no_defs")


#Old mask by me
atla <- atlantis_box %>%  st_transform(4326)
#WGOA <- st_crop(atla
#                    , xmin = -170, xmax = -147, ymin = 50, ymax = 60)
#WGOAaea <- st_transform(WGOA, crs = "+proj=aea +lat_0=50 +lon_0=-154 +lat_1=55 +lat_2=65 +x_0=0 +y_0=0 +ellps=GRS80 +units=m +no_defs")
#
#EGOA <- st_crop(atla
#                , xmin = -147, xmax = -125, ymin = 50, ymax = 60)
#
#EGOAaea <- st_transform(EGOA, crs = "+proj=aea +lat_0=50 +lon_0=-154 +lat_1=55 +lat_2=65 +x_0=0 +y_0=0 +ellps=GRS80 +units=m +no_defs")
#

```

## Area Type_ All Areas

```{r AK_ESR3, warning = FALSE, out.width="100%"}
akplots_area_type <- ggplot()+
  geom_sf(data=AKmngtA, aes(fill= Area_Type), color= NA, alpha=0.4)+
  theme_bw() +
  theme(panel.grid.major = element_line(color = "#cccccc", 
                                        linetype = "dashed", size = 0.5))+
  scale_fill_manual(values= c("ADFG Stat Area"="#999999", 
                              "BSIERP Region"="#56B4E9", 
                              "Ecosystem Area"="#009E73", 
                              "Ecosystem Subarea"="#F0E442", 
                              "Ecosystem Area Inside"="#0072B2", 
                              "Ecosystem Subarea Inside"="#D55E00", 
                              "NMFS Reporting Area"="#CC79A7"))+
  coord_sf()

akplots_area_type 

```

## Ecosystem Areas

```{r AK_ESR4, warning = FALSE, out.width="100%"}
akplots_Eco_Area <- ggplot()+
  geom_sf(data=AKmngtA, aes(fill= Ecosystem_Area), color= NA, alpha=0.3)+
  theme_bw() +
  theme(panel.grid.major = element_line(color = "#cccccc", 
                                        linetype = "dashed", size = 0.5))+
  scale_fill_manual(values= c("NA"="#999999", 
                              "Aleutian Islands"="#56B4E9", 
                              "Arctic"="#009E73", 
                              "Eastern Bering Sea"="#F0E442", 
                              "Gulf of Alaska"="#0072B2"))+
  coord_sf()

akplots_Eco_Area

```

## NMFS Areas

```{r AK_ESR5, warning = FALSE, out.width="100%"}
akplots_NMFS_Area <- AKmngtA %>% 
  drop_na(NMFS_REP_AREA) %>% 
  ggplot()+
  geom_sf(aes(color= NMFS_REP_AREA, geometry= geometry), fill="#999999", color= "#757575", alpha=0.8)+
  theme_bw() +
  theme(panel.grid.major = element_line(color = "#cccccc", 
                                        linetype = "dashed", size = 0.5))+
  geom_sf_text(
    aes(label = NMFS_REP_AREA),
    size = 2,
    hjust = 0.5, 
    color= "white"
  )+
  xlab("") + 
  ylab("") +
  coord_sf()

akplots_NMFS_Area
```


```{r BS_CS_EwEs, warning= FALSE, echo=FALSE}


#Chukchi <- read_sf("bering-sea-spatial-main_NOAA/shapefiles/Chukchi_Conner2019.shp")
#Chukchi <- st_transform(x = Chukchi, "+proj=aea +lat_0=50 +lon_0=-154 +lat_1=55 +lat_2=65 +x_0=0 +y_0=0 +ellps=GRS80 +units=m +no_defs")

NBS <- read_sf("bering-sea-spatial-main_NOAA/shapefiles/nebs_strata.shp")
NBS <- st_transform(x = NBS, "+proj=aea +lat_0=50 +lon_0=-154 +lat_1=55 +lat_2=65 +x_0=0 +y_0=0 +ellps=GRS80 +units=m +no_defs")

EBS <- read_sf("bering-sea-spatial-main_NOAA/shapefiles/mips.shp")
EBS <- st_transform(x = EBS, "+proj=aea +lat_0=50 +lon_0=-154 +lat_1=55 +lat_2=65 +x_0=0 +y_0=0 +ellps=GRS80 +units=m +no_defs")

EGOA <- read_sf("Ecopath_models_mask_Albi/EGOA.shp")
EGOA <- st_transform(x = EGOA, "+proj=aea +lat_0=50 +lon_0=-154 +lat_1=55 +lat_2=65 +x_0=0 +y_0=0 +ellps=GRS80 +units=m +no_defs")

WGOA <- read_sf("Ecopath_models_mask_Albi/WGOA.shp")
WGOA <- st_transform(x = WGOA, "+proj=aea +lat_0=50 +lon_0=-154 +lat_1=55 +lat_2=65 +x_0=0 +y_0=0 +ellps=GRS80 +units=m +no_defs")


```

```{r AK_ESR6, echo=FALSE, warning= FALSE}


### PWS - Eastern
PWS_E <- read_sf("PWS_Shapefiles_Bia/PWSeastern_Project.shp")
PWS_E <- st_transform(x = PWS_E, "+proj=aea +lat_0=50 +lon_0=-154 +lat_1=55 +lat_2=65 +x_0=0 +y_0=0 +ellps=GRS80 +units=m +no_defs")
#plot(PWS_E)


### PWS - Western
PWS_W <- read_sf("PWS_Shapefiles_BIA/PWSwestern_Project.shp")
PWS_W <- st_transform(x = PWS_W, "+proj=aea +lat_0=50 +lon_0=-154 +lat_1=55 +lat_2=65 +x_0=0 +y_0=0 +ellps=GRS80 +units=m +no_defs")
#plot(PWS_W)

# PWS 
PWS <- dplyr::bind_rows(list(PWS_E,PWS_W))

#Old central goa mask
#CGOAbd <- drop_na(AKmngtA, NMFS_REP_AREA)
#CGOAbd_v2 <-  CGOAbd[CGOAbd$NMFS_REP_AREA == c(630,640), ]
##CGOAbd_v3 <- CGOA_v2%>%  st_transform(4326) #to know the boundaries of 630 and 640 and crop from atlantis
#CGOA <- st_crop(atla
#                , xmin = -154.2662, xmax = -139.9997, ymin = 50, ymax = 60)

CGOA <- read_sf("Ecopath_models_mask_Albi/CGOA.shp")
CGOA_v2 <- st_transform(CGOA, crs = "+proj=aea +lat_0=50 +lon_0=-154 +lat_1=55 +lat_2=65 +x_0=0 +y_0=0 +ellps=GRS80 +units=m +no_defs")

```

## PWS model and LTER stations

```{r PWS_LTER, out.width="100%", warning=FALSE}

lter <- st_read("LTER_stations_TK.csv", options=c("longitude=x","latitude=y"))
#%>%filter(str_detect(station, 'PWS|IB|HB|MS|CG|1B|KI'))
lter <- st_as_sf(lter, coords = c("longitude", "latitude"), crs = 4326)

kerri <- st_read("C:/Users/biads/Dropbox/A_GWA/A_PWS_PAPER/PWS_EwE/DATA/Kerri_data_LTER/PWS_Chl_1997-2021_stations.csv", options=c("longitude=Longitude","latitude=Latitude"))%>%
  filter(str_detect(Station, 'KIP1|KIP2|KIP3|MS4|MS3|MS2|MS1|PWSW2|PWSW3|PWSW4|PWSW5|PWSW6|PWSW7|HB1|HB2|HB3|HB4|HB5|NI4|NI3|NI2|NI1|PWSW1|PWS11|PWS10|PWS9|PWS8|PWS7|PWS6|PWS5|PWS4|PWS3|PWS2|PWS1|MH1|MH2|MH3|MH4|PWS-2|PWSW3|CG2|PWSA|PWSB|PWSC|PWSE|IB0|IB1|IB2|CG1|KIP0|PWSF|PWSD|CG0|KIP2Prod|KIP2|PWS1|PWS2|KIP2|HB3|PSWE|PSW2'))
kerri <- st_as_sf(kerri, coords = c("Longitude", "Latitude"), crs = 4326)


#nmfs <- st_read("Ecopath_models_mask_Albi/data/shapefiles/gf95_nmfs/gf95_nmfs.shp")
nmfs <- AKmngtA %>% filter(NMFS_REP_AREA %in% c(620,630,640,649)) %>% # subset to 610-650
  dplyr::select(NMFS_REP_AREA, geometry)
#Using the NMFS bbox
#coast <- map("worldHires", regions = "USA", plot = FALSE, fill = TRUE)
#coast <- coast %>% st_as_sf() %>% st_transform(crs = st_crs(nmfs)) %>% st_crop(st_bbox(nmfs))
#coast <- coast %>%  st_transform(4326)

PWS_box <- PWS %>%  st_transform(4326)
#Using PWS bbox
AKpoly2 <- AKpoly %>% st_as_sf() %>% st_transform(crs = st_crs(PWS)) %>% st_crop(st_bbox(PWS))


PWS_plot <- AKpoly2 %>% 
  ggplot()+
  geom_sf(fill="gray", color="gray", alpha=0.5)+
  geom_sf(data=PWS, aes(fill="PWS"), color= NA, alpha=0.8)+
  geom_sf(data= lter)+
  #geom_sf(data= kerri)+
  theme_bw() +
  theme(panel.grid.major = element_line(color = "#cccccc", 
                                        linetype = "dashed", size = 0.5))+
  geom_sf_text(data= drop_na(lter, station),
               aes(label = station),
    size = 2,
    hjust = 1,
    vjust=1.5,
    color= "red")+
 # geom_sf_text(data= drop_na(kerri, Station),
 #   aes(label = Station),
 #   size = 2,
 #   hjust = 1,
 #   vjust=1.5,
 #   color= "black")+
  xlab("") + 
  ylab("") +
  coord_sf()

```

## LTER stations

```{r LTER, out.width="100%", warning=FALSE}


sfc_as_cols <- function(x, names = c("x","y")) {
  stopifnot(inherits(x,"sf") && inherits(sf::st_geometry(x),"sfc_POINT"))
  ret <- sf::st_coordinates(x)
  ret <- tibble::as_tibble(ret)
  stopifnot(length(names) == ncol(ret))
  x <- x[ , !names(x) %in% names]
  ret <- setNames(ret,names)
  dplyr::bind_cols(x,ret)
}

lter_all <- st_read("LTER_stations_TK.csv", options=c("longitude=x","latitude=y"))
lter_all <- st_as_sf(lter_all, coords = c("longitude", "latitude"), crs = 4326)

lter_all2 <- sfc_as_cols(lter_all)

#nmfs <- st_read("Ecopath_models_mask_Albi/data/shapefiles/gf95_nmfs/gf95_nmfs.shp")
nmfs <- AKmngtA %>% filter(NMFS_REP_AREA %in% c(630,640,649)) %>% # subset to 610-650
  select(NMFS_REP_AREA, geometry)

#Using PWS bbox
AKpoly3 <- AKpoly %>% st_as_sf() %>% st_transform(crs = st_crs(nmfs)) %>% st_crop(st_bbox(nmfs))


lter_plot <- AKpoly3 %>% 
  ggplot()+
  geom_sf(fill="gray", color="gray", alpha=0.5)+
  #geom_sf(data=PWS, aes(fill="PWS"), color= NA, alpha=0.8)+
  geom_sf(data= lter_all2)+
  theme_bw() +
  theme(panel.grid.major = element_line(color = "#cccccc", 
                                        linetype = "dashed", size = 0.5))+
  geom_sf_text(data= drop_na(lter_all2, station),
    aes(label = station),
    size = 2,
    hjust = 1,
    vjust=1.5,
    color= "black")+
  xlab("") + 
  ylab("") +
  coord_sf()

lter_plot+ geom_text_repel(data=lter_all2, aes(x= x, y=y, label=station))

```


## Ecosystem model and NMFS areas

```{r AK_plot_result, out.width="100%", warning=FALSE}
library(magick)

nmfs <- AKmngtA %>% filter(NMFS_REP_AREA %in% c(610,620,630,640, 650)) %>% # subset to 610-650
  dplyr::select(NMFS_REP_AREA, geometry)


akplots_v2 <- AKmngtA %>% 
  drop_na(Ecosystem_Area) %>% 
  ggplot()+
  geom_sf(data= drop_na(nmfs),aes(color= NMFS_REP_AREA, geometry= geometry), fill=NA, color= "#999999", alpha=0.8)+
  geom_sf(data=AKpoly,  fill= "#999999", color= NA, alpha=0.5)+
  geom_sf_text(data= drop_na(nmfs, NMFS_REP_AREA),
    aes(label = NMFS_REP_AREA),
    size = 2,
    hjust = 0.5, 
    color= "#333333"
  )+
  #geom_sf(aes(fill= Ecosystem_Area), color= "#888888", alpha=0.2)+
  #geom_sf(data=atlantis_box2, aes(fill="Atlantis"), color= NA, alpha=0.8)+
  #geom_sf(data=PWS, aes(fill="PWS"), color= NA, alpha=0.8)+
  #geom_sf(data=CGOA_v2, aes(fill= "NGOA"), color= NA, alpha=0.9)+
  #geom_sf(data=Chukchi, aes(fill= "Chukchi"), color= NA, alpha=0.5)+
  geom_sf(data=EBS, aes(fill= "EBS"), color= NA)+
  geom_sf(data=WGOA, aes(fill= "WGOA"), color= NA, alpha=0.75)+
  geom_sf(data=EGOA, aes(fill= "EGOA"), color= NA, alpha=0.75)+
  geom_sf(data=NBS, aes(fill= "NBS"), color= NA)+
  #geom_sf(data= lter)+
  #geom_sf(data= kerri)+
  #geom_sf_pattern(data=WGOA, aes(pattern_type="WGOA"), pattern = 'magick',pattern_fill = "#E69F00",  pattern_alpha=0.6, color="#E69F00", fill=NA)+
  #geom_sf_pattern(data=EGOA, aes(pattern_type="EGOA"), pattern = 'magick',pattern_density= 0.01, pattern_fill = "#E69F00", pattern_alpha=0.6, color="#E69F00", fill=NA)+
    theme_bw() +
  theme(panel.grid.major = element_line(color = "#cccccc", 
                                        linetype = "dashed", size = 0.5))+
  #geom_sf_text(data= drop_na(lter, station),
  #             aes(label = station),
  #  size = 2,
  #  hjust = 1,
  #  vjust=1.5,
  #  color= "red")+
  theme_bw() +
  theme(panel.grid.major = element_line(color = "#cccccc", 
                                        linetype = "dashed", size = 0.5))+
  xlab("") + 
  ylab("") +
  scale_fill_manual(values= c( "WGOA"= "#E69F00",
                              #"PWS"="#D55E00",
                              "EGOA"= "#009E73",
                              #"EGOA" ="#999933",
                              #"Chukchi"="#0072B2", 
                              "NBS"="#0072B2", 
                              "EBS"="#56B4E9" 
                              #"Central_GOA" = "#CC79A7",
                              #"Eastern Bering Sea"="#F0E442", 
                              #"Gulf of Alaska"="#0072B2"
                              ))+
  #scale_pattern_type_manual(values = c("WGOA"="left30", 
  #                                     "EGOA"="right45")) +
   labs( x = "", y = "", fill = " ", pattern_type=" ") +
  
  coord_sf()
  
  
akplots_v2
ggsave( "FIGURES/EWE_ak_plots.png", akplots_v2, height=4, width=8, dpi = 600)
```

## Ecosystem model, LTER and NMFS areas for PWS

```{r AK_plot_result_v2, out.width="100%", warning=FALSE}

nmfs <- AKmngtA %>% filter(NMFS_REP_AREA %in% c(610,620,630,640,649, 650)) %>% # subset to 610-650
  dplyr::select(NMFS_REP_AREA, geometry)


akplots_v3 <- AKmngtA %>% 
  drop_na(Ecosystem_Area) %>% 
  ggplot()+
  geom_sf(data= drop_na(nmfs),aes(color= NMFS_REP_AREA, geometry= geometry), fill=NA, color= "grey", alpha=0.8)+
  geom_sf(data=atlantis_box2, aes(fill="Atlantis"), color= NA, alpha=0.8)+ 
  geom_sf(data=PWS, aes(fill="PWS"), color= NA, alpha=0.8)+
  geom_sf(data=CGOA_v2, aes(fill= "NGOA"), color= NA, alpha=0.9)+
  geom_sf(data= lter, size=0.5)+
  geom_sf_pattern(data=WGOA, aes(pattern_type="WGOA"), pattern = 'magick',pattern_fill = "#E69F00",  pattern_alpha=0.6, color="#E69F00", fill=NA)+
  geom_sf_pattern(data=EGOA, aes(pattern_type="EGOA"), pattern = 'magick',pattern_density= 0.01, pattern_fill = "#E69F00", pattern_alpha=0.6, color="#E69F00", fill=NA)+
  #geom_sf(data=AKpoly,  fill= "#999999", color= NA, alpha=0.5)+
  geom_sf_text(data= drop_na(nmfs),
    aes(label = NMFS_REP_AREA),
    size = 3,
    hjust = 0.5, 
    color= "black"
  )+
    theme_bw() +
  theme(panel.grid.major = element_line(color = "#cccccc", 
                                        linetype = "dashed", size = 0.5))+
  theme_bw() +
  theme(panel.grid.major = element_line(color = "#cccccc", 
                                    linetype = "dashed", size = 0.5))+
  xlab("") + 
  ylab("") +
  scale_fill_manual(values= c( "PWS"= "#D55E00",
                              "Atlantis"="#E69F00",
                              "NGOA"= "#009E73"))+
   scale_pattern_type_manual(values = c("WGOA"="left30", 
                                        "EGOA"="right45")) +
   labs( x = "", y = "", fill = " ", pattern_type=" ") +
  
  coord_sf()
  
  
akplots_v3
```

## Survey responses

## What is your main goal and what do you aim to achieve with your modeling approach? (e.g., model purpose, model hypothesis).

- Bia: 
    1. How marine heatwaves affect the marine ecosystem resistance and resilience. 
    2. In addition my main question, I want to explore what takes to recover Pacific herring populations.
- Albi:
    1. To model ecosystem-level effects of past and future climate events
- Andy: 
    1. Evaluate structure/function of the food web, gap analysis, resilience to perturbations


## What is your main goal. cnt.

  - Bridget:
    1. Spatially align WCGOA Ecopath model to GOA-CLIM for future model comparison studies
    2. Update biomass, catch, ROMs 10km (temperature, phytoplankton, zooplankton) time series to perform hindcasts/forecasts on environmental relationships and food web dynamics through time including marine heatwave time periods (2014-2019)
    3. Add temperature functions (consumption and respiration) to address temperature dependent changes in food web dynamics
    4. Inform groundfish Fisheries Management Process through the Ecosystem Status Report by providing modeled syntheses and indicators of food web structure and dynamics 
    5. Automate model updates & Accessible & Applied (Github, etc)
    6. Set up model for future model comparison/ multi-model approaches through GOA-CLIM 
    
## What is your main goal. cnt.

  - Szymon
    1. quantify and analyze the structure of the eastern Gulf of Alaska ecosystem
    2. hindcast ecosystem dynamics in the eastern Gulf of Alaska iii) forecast these ecosystem dynamics under several climate change and fisheries management scenarios
    3. compare ecosystem structure and dynamics in the eastern vs. central and western Gulf of Alaska ecosystem (Bridget's model) and in Ecopath vs. Atlantis (Albi's model).


## Are you using any forcing function? If so, what are the forcing functions being used? 

- Bia: 
    DFA generated forcing functions
- Bridget:
    1. Temperature
    2. Phytoplankton
    3. Catch
    4. Diet maybe future"
- Albi:
    (interpreting "forcing" in an Atlantis context) ROMS daily data for temperature, salinity, water transport, solar radiation. Catch from AKFIN.
- Szymon
    I will be using a temperature time series to allow species to respond to climate conditions according to their thermal niche parameters.

